---
title: "Build and deploy a stroke prediction model using R"
date: "`r Sys.Date()`"
output: html_document
author: "Put your name!"
---

# About Data Analysis Report

This RMarkdown file contains the report of the data analysis done for the project on building and deploying a stroke prediction model in R. It contains analysis such as data exploration, summary statistics and building the prediction models. The final report was completed on `r date()`. 

**Data Description:**

According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

This data set is used to predict whether a patient is likely to get stroke based on the input parameters like gender, age, various diseases, and smoking status. Each row in the data provides relevant information about the patient.


# Task One: Import data and data preprocessing

## Load data and install packages

```{r}
# Install packages if needed
# install.packages("tidyverse")
# install.packages("corrplot")
# install.packages("mice")
# install.packages("DataExplorer")
# install.packages("janitor")
# install.packages("skimr")

# Load packages
library(tidyverse)
library(skimr)
library(janitor)
library(corrplot)
library(mice)
library(DataExplorer)


stroke <- read.csv("healthcare-dataset-stroke-data.csv")

print(head(stroke))

print(dim(stroke))

print(str(stroke))
print(summary(stroke))

colSums(is.na(stroke))

#Remove ID Column-The ID column does not contribute to prediction.
stroke <- stroke[, !(names(stroke) %in% c("id"))]

#Convert Variables to Appropriate Data Types
stroke <- stroke %>%
  mutate(
    gender = as.factor(gender),
    hypertension = as.factor(hypertension),
    heart_disease = as.factor(heart_disease),
    ever_married = as.factor(ever_married),
    work_type = as.factor(work_type),
    Residence_type = as.factor(Residence_type),
    smoking_status = as.factor(smoking_status),
    stroke = as.factor(stroke)
)

print(str(stroke))

#Fix BMI Data Type
stroke$bmi[stroke$bmi == "N/A"] <- NA
stroke$bmi <- as.numeric(stroke$bmi)


#4.6 Handle Inconsistent Values
unique(stroke$gender)
unique(stroke$work_type)
unique(stroke$Residence_type)
unique(stroke$smoking_status)

stroke$smoking_status[stroke$smoking_status == "Unknown"] <- NA

#Check Duplicate Records
sum(duplicated(stroke))
stroke <- distinct(stroke)


```




## Describe and explore the data


```{r}
#Summary Statistics
summary(stroke)

#5.2 Distribution of Stroke Cases
library(ggplot2)

ggplot(stroke,aes(x=stroke,fill = stroke))+
  geom_bar() + 
  labs(title = "Distribution of Stroke Cases",x = "Stroke",y = "Count")+
  theme_minimal()


#Age Distribution
ggplot(stroke,aes(x= age))+
  geom_histogram(bins = 30)+
  labs(title = "Age Distribution",
    x = "Age",
    y = "Frequency")+
  theme_minimal()

#Age and Stroke Relationship(theme_minmal - remove the gray color background in the charts)
ggplot(stroke,aes(x=stroke,y=age,fill=stroke))+
  geom_boxplot()+
  labs(title = "Age by Stroke Status",
    x = "Stroke",
    y = "Age")+
  theme_minimal()


#Hypertension and Stroke(fill- fill the colors for bars, position ="fill" - it show like presentage wise)
ggplot(stroke,aes(x=hypertension,fill=stroke))+
  geom_bar(position = "fill")+
  labs(title = "Stroke by Hypertension Status",
    x = "Hypertension",
    y = "Proportion")+
  theme_minimal()

#Heart Disease and Stroke
ggplot(stroke,aes(x=heart_disease,fill=stroke))+
  geom_bar(position = "fill")+
  labs(title = "Stroke by Heart Disease Status",
    x = "Heart Disease",
    y = "Proportion")+
  theme_minimal()

#Average Glucose Level Distribution
ggplot(stroke,aes(x=avg_glucose_level))+
  geom_histogram(bins = 30)+
  labs(tittle="Avg Glucose",x="avg_glucose")+
  theme_minimal()

#Average Glucose Level by Stroke Status
ggplot(stroke,aes(x=stroke,y=avg_glucose_level,fill=stroke))+
  geom_boxplot()+
  labs(title = "Average Glucose Level by Stroke Status",
    x = "Stroke",
    y = "Average Glucose Level")+
  theme_minimal()

#BMI Distribution
ggplot(stroke,
       aes(x = bmi)) +
  geom_histogram(bins = 30) +
  labs(
    title = "BMI Distribution",
    x = "BMI",
    y = "Frequency"
  ) +
  theme_minimal()

#Smoking Status and Stroke
ggplot(stroke,
       aes(x = smoking_status,
           fill = stroke)) +
  geom_bar(position = "fill") +
  labs(
    title = "Stroke by Smoking Status",
    x = "Smoking Status",
    y = "Proportion"
  ) +
  theme_minimal()

#Gender and Stroke
ggplot(stroke,
       aes(x = gender,
           fill = stroke)) +
  geom_bar(position = "fill") +
  labs(
    title = "Stroke by Gender",
    x = "Gender",
    y = "Proportion"
  ) +
  theme_minimal()


#Correlation Analysis
install.packages("corrplot")
library(corrplot)

#Select numeric variables then store in to new variable
numeric_vars <- stroke[, c("age",
                           "avg_glucose_level",
                           "bmi")]

#cor() - Calculates the correlation between variables.A correlation measures how strongly two variables move together.(+1 positive,0 no relationship,-1 negative relationship)
#use = "complete.obs"- Ignore rows that contain missing values (NA)
cor_matrix <- cor(
  numeric_vars,
  use = "complete.obs"
)
print(cor_matrix)

#Visualize the correlation matrix -Draws a graphical representation of the correlations.
#method = "circle"-Uses circles to represent correlations.(Larger circle = stronger correlation,Smaller circle = weaker correlation)
corrplot(
  cor_matrix,
  method = "circle"
)


#(Key Insights from EDA
#The dataset is highly imbalanced, with relatively few stroke cases.
#Age appears to be one of the strongest predictors of stroke.
#Patients with hypertension show a higher prevalence of stroke.
#Heart disease is positively associated with stroke occurrence.
#Higher average glucose levels are observed among stroke patients.
#Smoking status may contribute to stroke risk.
#BMI shows some variation across patients but requires missing value imputation.
#Several variables demonstrate predictive potential and should be retained for model development.)**

```





# Task Two: Build prediction models

```{r}
#Introduction to Tidymodels-Tidymodels is a collection of R packages designed for machine learning and data modeling using a consistent and user-friendly workflow. It follows the principles of the tidyverse, making data science tasks easier to read, write, and maintain.Instead of learning different functions for each machine learning algorithm, tidymodels provides a unified framework forData preprocessing Splitting data,Building models,Training models,Evaluating performance,Tuning hyperparameters
# install.packages("tidymodels")
# install.packages("themis")
# install.packages("vip")
# install.packages("xgboost")
# install.packages("ranger")
library(tidymodels)
library(themis)      # SMOTE
library(vip)
library(xgboost)
library(ranger)

#Train-Test Split
#set.seed(123) - Fixes randomness in R.When you split data, R uses random sampling.
#initial_split() - Split dataset into train/test
#stroke_clean - This is your cleaned dataset (no missing values or processed data).
#strata = stroke - Keep the same proportion of stroke cases in both train and test sets.
set.seed(123)

stroke_clean <- stroke
stroke_split <- initial_split(
  stroke_clean,
  prop = 0.8,
  strata = stroke
)

stroke_train <- training(stroke_split)
stroke_test <- testing(stroke_split)

table(stroke_train$stroke)
table(stroke_test$stroke)


#Create Preprocessing Recipe
#stroke = target variable (what you want to predict), "." = all other columns are predictors (features),stroke_train = training dataset
#👉 “Use all variables to predict stroke using the training data.”
stroke_recipe <- recipe(
  stroke ~ .,
  data = stroke_train
)%>%
  step_impute_median(all_numeric_predictors()) %>% #Handle missing numeric values - Finds missing values in numeric columns,Replaces them with the median value(AGE - Median = 52.5 → replaces NA with 52.5)
  step_impute_mode(all_nominal_predictors()) %>% #Handle missing categorical values - Fills missing categorical values,Uses the most frequent category (mode) - (smoked_status Mode = "never smoked" → fills NA)
  step_dummy(all_nominal_predictors()) %>% #Convert categorical variables into numbers - Machine learning models need numbers, not text.(male -> 1female -> 0)
  step_normalize(all_numeric_predictors()) %>% #Normalize numeric values - Scales numeric variables to same range
  step_smote(stroke) #Handle class imbalance - SMOTE = Synthetic Minority Over-sampling Technique(0(No stroke)	High,1(Stroke)	Low),  Creating synthetic (fake but realistic) stroke cases,Balancing the dataset, Without this:Model learns mostly "No Stroke",Poor prediction for stroke cases



stroke_recipe

#Cross Validation - It splits your training data into multiple parts for model testing.(Split my training data into 10 parts so I can test my model 10 different times in a fair way.) Benift : Avoid overfitting,Get more reliable accuracy,Test model on multiple splits,Compare models fairly

#You divide your training data into 10 equal parts (folds):Fold 1Fold 2Fold 3...Fold 10
#v = 10 “Split data into 10 folds”
#strata = stroke Keeps the same proportion of stroke cases in each fold (Without stratification:Some folds may have no stroke cases,Model evaluation becomes wrong   With stratification:Each fold has same stroke ratio)

set.seed(123)
cv_folds <-vfold_cv(stroke_train, v = 10, strata = stroke)


#Goal - You want to build a model that predicts:Will a person have a stroke?  0 = No Stroke,1 = Stroke
#Model 1: Logistic Regression

#Step 1: Create the Logistic Regression Model - "I want to use Logistic Regression for a classification problem."
#logistic_reg() - Create an empty logistic regression model
#set_engine("glm") - Use R's built-in logistic regression engine(glm = Generalized Linear Model (built into R))
#set_mode("classification") - Tells R:I am predicting categories.Not numbers.Because:Stroke = Yes or No.This is a classification problem.

log_model <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

#Step 2: Create a Workflow - Data Preparation + Model into one object.
#workflow() - Create an empty machine learning pipeline
#add_recipe(stroke_recipe) - Adds your preprocessing steps.(Fill missing values,Convert text to numbers,Normalize data,Apply SMOTE)
#add_model(log_model) - Adds the logistic regression model.

log_wf <- workflow() %>%
  add_recipe(stroke_recipe) %>%
  add_model(log_model)

#Step 3: Train and Test the Model
#fit_resamples() - Train the model many times using cross-validation
#resamples = cv_folds - Uses the 10 folds you created earlier.(Run 1:Train on 9 folds,Test on 1 fold .Run 2:Train on different 9 folds,Test on another fold....Run 10:Train on 9 folds,Test on last fold.....After 10 runs:Average all results This gives a more reliable evaluation.)
#metrics = metric_set(...)

#accuracy - How many predictions were correct?(Example:100 people tested,95 predicted correctly..Accuracy = 95%)
#roc_auc - How well the model separates Stroke and No Stroke ,Values:0.5 = Bad,1.0 = Perfect...Higher is better.
#sens(Sensitivity)- How many actual stroke cases did the model find?Example:100 stroke patients,80 detected,...Sensitivity = 80%
#spec(Specificity) - How many non-stroke patients were correctly identified?Example:100 non-stroke patients,95 correctly identified...Specificity = 95%

log_fit <- fit_resamples(
  log_wf,
  resamples = cv_folds,
  metrics = metric_set(
    accuracy,
    roc_auc,
    sens,
    spec
  )
)

collect_metrics(log_fit)



#Model 2: Decision Tree - A Decision Tree makes decisions using a series of questions.Example:Age > 60?Yes.. -> Hypertension? Yes → Stroke,No  → No Stroke....It looks like a tree with branches.

#Step 1: Create the Decision Tree Model
#decision_tree() - Create a Decision Tree model
#cost_complexity = tune() - This controls tree pruning.(tune() = Try different values and choose the best one)
#tree_depth = tune() - Controls how deep the tree can grow.(example -Depth = 2 Age? ->yes,no...depth = 10 age?,BMI?,Gloucose?....)   tune() = Find the best depth automatically
#set_engine("rpart") - Use the rpart package to build the decision tree
#

tree_model <- decision_tree(
  cost_complexity = tune(),
  tree_depth = tune()
) %>%
  set_engine("rpart") %>%
  set_mode("classification")

#Step 2: Create a Workflow - Combine:Recipe + Decision Tree

tree_wf <- workflow() %>%
  add_recipe(stroke_recipe) %>%
  add_model(tree_model)

#Step 3: Create a Grid of Values - We don't know the best: cost_complexity, tree_depth So we test many combinations.
#levels = 5- Try 5 different values for each parameter

tree_grid <- grid_regular(
  cost_complexity(),
  tree_depth(),
  levels = 5
)

#Step 4: Tune the Model
#tune_grid() - Try all parameter combinations and find the best one
#



tree_tuned <- tune_grid(
  tree_wf,
  resamples = cv_folds,
  grid = tree_grid,
  metrics = metric_set(
    accuracy,
    roc_auc
  )
)



#Model 3: Random Forest - A Random Forest is a collection of many Decision Trees.This usually gives better accuracy than a single Decision Tree.

#Step 1: Create the Random Forest Model
#rand_forest() - Create a forest of decision trees
#trees = 500 - Build 500 decision trees(Then combine their predictions.More trees generally make the model more stable.)
#mtry = tune() - When a tree splits data, it chooses from a subset of variables.(age,bmi,glucose,hypertension,heart_disease) Instead of checking all variables, Random Forest randomly checks only a few."How many predictors can be considered at each split?"
#min_n = tune() - Minimum number of observations in a leaf node(Leaf A = 50 people,Leaf B = 30 people,Leaf C = 2 people)
#set_engine("ranger") - Use the ranger package to build the Random Forest

rf_model <- rand_forest(
  trees = 500,
  mtry = tune(),
  min_n = tune()
) %>%
  set_engine("ranger") %>%
  set_mode("classification")

#Step 2: Create a Workflow(Recipe+Random Forest)

rf_wf <- workflow() %>%
  add_recipe(stroke_recipe) %>%
  add_model(rf_model)

#Step 3: Create a Tuning Grid - We don't know the best:mtry,min_n ...So we test different values.
#mtry(range = c(2,10)) - ry values between:2 and 10 For example:2 4 6 8 10
#min_n() - Try different values for minimum leaf size.(2,5,10,15,20)
#levels = 5 - Use 5 values for each parameter
#

rf_grid <- grid_regular(
  mtry(range = c(2,10)),
  min_n(),
  levels = 5
)


#Step 4: Tune the Model
#tune_grid() -Try all parameter combinations and find the best one
#grid = rf_grid - Uses all combinations from your tuning grid.
#
rf_tuned <- tune_grid(
  rf_wf,
  resamples = cv_folds,
  grid = rf_grid,
  metrics = metric_set(
    accuracy,
    roc_auc
  )
)



#Model 4: Support Vector Machine - An SVM is a machine learning algorithm that tries to find the best boundary between classes.The SVM tries to find the boundary that separates the two groups as clearly as possible.

#Step 1: Create the SVM Model
#svm_rbf() - Creates a Support Vector Machine with an RBF kernel.The RBF kernel allows the SVM to create curved boundaries.
#cost = tune() - Cost controls how strict the SVM is.(Small cost:More flexible,Allows some mistakes...Large cost:Very strict,Tries to classify everything correctly)
#rbf_sigma = tune() - This controls the shape of the curved boundary.(Small sigma:Smooth boundary...Large sigma:Complex boundary)
#set_engine("kernlab") - Use the kernlab package to build the SVM


install.packages("kernlab")
library(kernlab)
svm_model <- svm_rbf(
  cost = tune(),
  rbf_sigma = tune()
) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

#Step 2: Create a Workflow(Recipe+ SVM Model)

svm_wf <- workflow() %>%
  add_recipe(stroke_recipe) %>%
  add_model(svm_model)

#Step 3: Tune the SVM
#tune_grid() - Try many combinations of cost and sigma
#grid = 10 - Generate 10 different combinations of cost and sigma
#


svm_tuned <- tune_grid(
  svm_wf,
  resamples = cv_folds,
  grid = 10,
  metrics = metric_set(
    accuracy,
    roc_auc
  )
)




#Model 5: XGBoost - Instead of creating one tree or many independent trees, it builds trees one after anotherExample : Tree 1:Predicts 80% correctly -> Tree 2:Focuses on Tree 1's mistakes -> Tree 3:Focuses on remaining mistakesAnd so on...Tree 1,Tree 2 fixes mistakes,Tree 3 fixes mistakes,Tree 4 fixes mistakes...This often produces very high accuracy.)


#Step 1: Create the XGBoost Model
#boost_tree() - Create a boosted tree model(Creates an XGBoost model.)
#trees = 500 - Build 500 trees Each tree learns from errors made by previous trees.
#tree_depth = tune() - Controls how deep each tree can grow.
#learn_rate = tune() - It controls:How fast the model learns(Large learning rate-Learns quickly,May miss optimal solution....Small learning rate- Learns slowly,Often more accurate)
#min_n = tune() - Minimum observations require for a split
#set_engine("xgboost") - Use the xgboost package

xgb_model <- boost_tree(
  trees = 500,
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

#Step 2: Create a Workflow(Recipe+XGBoost Model)

xgb_wf <- workflow() %>%
  add_recipe(stroke_recipe) %>%
  add_model(xgb_model)

#Step 3: Tune the Model
#tune_grid() - Try different parameter combinations and find the best one
#grid = 10 - Generate 10 different combinations of parameters

xgb_tuned <- tune_grid(
  xgb_wf,
  resamples = cv_folds,
  grid = 10,
  metrics = metric_set(
    accuracy,
    roc_auc
  )
)


collect_metrics(log_fit)

collect_metrics(tree_tuned)

collect_metrics(rf_tuned)

#collect_metrics(svm_tuned)

collect_metrics(xgb_tuned)


```




# Task Three: Evaluate and select prediction models

```{r}
library(tidymodels)
library(yardstick)
# 
# best_rf <- select_best(rf_tuned, "roc_auc")
# final_rf_wf <- finalize_workflow(rf_wf, best_rf)
# rf_fit <- fit(final_rf_wf, stroke_train)
# 
# rf_pred_prob <- predict(rf_fit, stroke_test, type = "prob")
# rf_pred_class <- predict(rf_fit, stroke_test, type = "class")
# rf_results <- bind_cols(
#   stroke_test,
#   rf_pred_class,
#   rf_pred_prob
# )
# 
# log_fit <- fit(log_wf, stroke_train)
# 
# log_pred_class <- predict(log_fit, stroke_test, type = "class")
# log_pred_prob  <- predict(log_fit, stroke_test, type = "prob")
# 
# log_results <- bind_cols(
#   stroke_test,
#   log_pred_class,
#   log_pred_prob
# )
# 
# roc_auc(log_results,
#         truth = stroke,
#         .pred_1)
# 
# 
# 
# 
# 
# best_xgb <- select_best(xgb_tuned, "roc_auc")
# 
# final_xgb_wf <- finalize_workflow(xgb_wf, best_xgb)
# 
# xgb_fit <- fit(final_xgb_wf, stroke_train)
# 
# xgb_pred_class <- predict(xgb_fit, stroke_test, type = "class")
# xgb_pred_prob  <- predict(xgb_fit, stroke_test, type = "prob")
# 
# 
# 
# conf_mat(rf_results,
#          truth = stroke,
#          estimate = .pred_class)
# 
# 
# metric_set(accuracy, sens, spec, precision, f_meas)(
#   rf_results,
#   truth = stroke,
#   estimate = .pred_class
# )
# 
# 
# roc_auc(rf_results,
#         truth = stroke,
#         .pred_1)
# 
# 
# roc_curve(rf_results,
#           truth = stroke,
#           .pred_1) %>%
#   autoplot()
# 
# roc_auc(log_results, truth = stroke, .pred_1)
# roc_auc(rf_results, truth = stroke, .pred_1)
# roc_auc(svm_results, truth = stroke, .pred_1)
# roc_auc(xgb_results, truth = stroke, .pred_1)
# 

best_xgb <- select_best(xgb_tuned, metric = "roc_auc")

final_xgb_wf <- finalize_workflow(xgb_wf, best_xgb)

xgb_fit <- fit(final_xgb_wf, stroke_train)

xgb_pred_class <- predict(xgb_fit, stroke_test, type = "class")
xgb_pred_prob  <- predict(xgb_fit, stroke_test, type = "prob")

xgb_results <- stroke_test %>%
  bind_cols(xgb_pred_prob)

roc_auc(xgb_results, truth = stroke, .pred_1)

```



# Task Four: Deploy the prediction model

```{r}
# 
# install.packages("shinydashboard")
# install.packages("rmarkdown")
# install.packages("shiny")
# install.packages("plotly")



library(shiny)
library(shinydashboard)
library(plotly)
library(rmarkdown)

# =====================
# MODEL PLACEHOLDER
# =====================
# If you have your actual model saved as an .rds file, uncomment the line below:
# final_model <- readRDS("path_to_your_model.rds")

final_model <- xgb_fit
saveRDS(final_model, "model.rds")
getwd()

# =====================
# UI
# =====================
ui <- dashboardPage(
  skin = "black", 
  
  dashboardHeader(title = "Stroke Prediction"),
  
  dashboardSidebar(
    selectInput("gender", "Gender:", c("Male", "Female")),
    numericInput("age", "Age:", 65),
    
    selectInput("hypertension", "Hypertension:", c("No" = "0", "Yes" = "1"), selected = "1"),
    selectInput("heart_disease", "Heart Disease:", c("No" = "0", "Yes" = "1"), selected = "1"),
    
    selectInput("ever_married", "Are you married?", c("No", "Yes"), selected = "Yes"),
    
    selectInput("work_type", "Work Type:",
                c("children", "Govt_job", "Never_worked", "Private", "Self-employed"),
                selected = "Self-employed"),
                
    selectInput("Residence_type", "Residence Type:",
                c("Rural", "Urban"), selected = "Rural"),
    
    numericInput("glucose", "Average Glucose Level:", 100),
    numericInput("bmi", "BMI:", 25),
    
    selectInput("smoking_status", "Smoking Status:",
                c("never smoked", "formerly smoked", "smokes", "Unknown")),
    
    br(),
    div(style = "text-align: center;",
        actionButton("predict", "Predict", class = "btn-primary", style = "width: 80%;"),
        br(), br(),
        downloadButton("download_report", "Download PDF Report", class = "btn-success", style = "width: 80%;")
    )
  ),
  
  dashboardBody(
    # Dynamic display container that toggles based on user action
    uiOutput("dashboard_display")
  )
)

# =====================
# SERVER
# =====================
server <- function(input, output, session) {
  
  results <- reactiveValues(stroke_p = 0, no_stroke_p = 0)
  
  shap_values <- reactiveVal(data.frame(
    Feature = character(),
    Contribution = numeric()
  ))
  
  history_data <- reactiveVal(data.frame(
    `Patient ID` = integer(),
    Age = numeric(),
    Gender = character(),
    Hypertension = character(),
    `Heart Disease` = character(),
    `Glucose Level` = numeric(),
    BMI = numeric(),
    `Stroke Risk` = character(),
    check.names = FALSE,
    stringsAsFactors = FALSE
  ))
  
  # Track run counts; starts at 0 to keep the dashboard initial state blank
  counter <- reactiveVal(0)
  
  observeEvent(input$predict, {
    counter(counter() + 1)
    
    new_data <- data.frame(
      age = input$age,
      gender = input$gender,
      hypertension = input$hypertension,
      heart_disease = input$heart_disease,
      ever_married = input$ever_married,
      work_type = input$work_type,
      Residence_type = input$Residence_type,
      avg_glucose_level = input$glucose,
      bmi = input$bmi,
      smoking_status = input$smoking_status
    )
    
    # --- OPTION A: YOUR REAL ML MODEL & SHAP PIPELINE ---
    # pred_prob <- predict(final_model, new_data, type = "prob")
    # results$stroke_p <- round(pred_prob[[2]] * 100, 2)
    # results$no_stroke_p <- round(pred_prob[[1]] * 100, 2)
    
    # --- OPTION B: LIVE DEMO SIMULATION ---
    w_age <- round(input$age * 0.65, 2)
    w_glucose <- round(input$glucose * 0.12, 2)
    w_hyper <- ifelse(input$hypertension == "1", 15.00, 1.50)
    w_heart <- ifelse(input$heart_disease == "1", 18.00, 1.20)
    w_bmi <- round(input$bmi * 0.25, 2)
    
    base_risk <- w_age + w_glucose + w_hyper + w_heart
    simulated_stroke <- min(99.9, max(0.1, round(base_risk, 2)))
    
    results$stroke_p <- simulated_stroke
    results$no_stroke_p <- round(100 - simulated_stroke, 2)
    
    df_shap <- data.frame(
      Feature = c("Age", "Heart Disease", "Hypertension", "Glucose Level", "BMI"),
      Contribution = c(w_age, w_heart, w_hyper, w_glucose, w_bmi)
    )
    df_shap <- df_shap[order(df_shap$Contribution), ]
    shap_values(df_shap)
    
    new_row <- data.frame(
      `Patient ID` = paste0("#", 1000 + counter()),
      Age = input$age,
      Gender = input$gender,
      Hypertension = ifelse(input$hypertension == "1", "Yes", "No"),
      `Heart Disease` = ifelse(input$heart_disease == "1", "Yes", "No"),
      `Glucose Level` = input$glucose,
      BMI = input$bmi,
      `Stroke Risk` = paste0(results$stroke_p, "%"),
      check.names = FALSE,
      stringsAsFactors = FALSE
    )
    
    current_history <- history_data()
    updated_history <- rbind(new_row, current_history)
    history_data(updated_history)
  })
  
  # Dynamic UI Switch Manager
  output$dashboard_display <- renderUI({
    if (counter() == 0) {
      # FIRST IMPRESSION STATE: Empty placeholder panel with clean instructions
      fluidRow(
        box(
          title = "Patient Diagnostic System",
          width = 12,
          status = "primary",
          solidHeader = TRUE,
          p(style = "font-size: 16px; text-align: center; padding: 40px 0;",
            "No data analyzed yet. Please adjust the clinical parameters in the left sidebar menu and click ",
            tags$strong("Predict"), " to evaluate stroke risk profiles.")
        )
      )
    } else {
      # CALCULATED STATE: Shows complete analytical components cleanly
      tagList(
        fluidRow(
          box(
            title = "Prediction Summary", 
            width = 12, 
            collapsible = TRUE,
            uiOutput("prediction_cards")
          )
        ),
        fluidRow(
          box(
            title = "Risk Distribution Breakdown",
            width = 5,
            collapsible = TRUE,
            plotlyOutput("stroke_plot", height = "300px")
          ),
          box(
            title = "Patient-Specific Risk Drivers (SHAP Value Insights)",
            width = 7,
            collapsible = TRUE,
            plotlyOutput("importance_plot", height = "300px")
          )
        ),
        fluidRow(
          box(
            title = "Patient Risk History Log",
            width = 12,
            collapsible = TRUE,
            status = "primary",
            div(style = "overflow-x: auto;", 
                tableOutput("history_table")
            )
          )
        )
      )
    }
  })
  
  # Render custom colored UI metric cards
  output$prediction_cards <- renderUI({
    req(counter() > 0)
    fluidRow(
      column(width = 6,
             div(style = "background-color: #c94c3a; color: white; padding: 20px; border-radius: 3px; position: relative; margin-bottom: 10px;",
                 div(style = "font-size: 38px; font-weight: bold; line-height: 1.2;", paste0(results$stroke_p, "%")),
                 div(style = "font-size: 14px; font-weight: 500; opacity: 0.9;", "Probability of having a Stroke"),
                 div(class = "fa fa-info", style = "position: absolute; right: 25px; bottom: 15px; font-size: 55px; opacity: 0.15;")
             )
      ),
      column(width = 6,
             div(style = "background-color: #27ae60; color: white; padding: 20px; border-radius: 3px; position: relative; margin-bottom: 10px;",
                 div(style = "font-size: 38px; font-weight: bold; line-height: 1.2;", paste0(results$no_stroke_p, "%")),
                 div(style = "font-size: 14px; font-weight: 500; opacity: 0.9;", "Probability of not having a Stroke"),
                 div(class = "fa fa-info", style = "position: absolute; right: 25px; bottom: 15px; font-size: 55px; opacity: 0.15;")
             )
      )
    )
  })
  
  # Render Donut Chart
  output$stroke_plot <- renderPlotly({
    req(counter() > 0)
    plot_ly(
      labels = c("No Stroke Risk", "Stroke Risk"),
      values = c(results$no_stroke_p, results$stroke_p),
      type = 'pie',
      hole = 0.6,
      marker = list(colors = c("#27ae60", "#c94c3a")),
      textinfo = 'percent',
      hoverinfo = 'text',
      text = ~paste0(c("No Stroke Risk: ", "Stroke Risk: "), c(results$no_stroke_p, results$stroke_p), "%")
    ) %>%
      layout(
        showlegend = TRUE,
        legend = list(orientation = "h", xanchor = "center", x = 0.5, y = -0.1),
        margin = list(l = 10, r = 10, t = 10, b = 10)
      )
  })
  
  # Render SHAP Chart
  output$importance_plot <- renderPlotly({
    req(counter() > 0)
    df <- shap_values()
    df$Feature <- factor(df$Feature, levels = df$Feature)
    
    plot_ly(
      data = df,
      x = ~Contribution,
      y = ~Feature,
      type = 'bar',
      orientation = 'h',
      marker = list(color = '#34495e'),
      hoverinfo = 'text',
      text = ~paste0("Risk Weight Contribution: +", Contribution)
    ) %>%
      layout(
        xaxis = list(title = "Relative Risk Impact Weight"),
        yaxis = list(title = ""),
        margin = list(l = 100, r = 20, t = 20, b = 40)
      )
  })
  
  # Render Patient History Log Table
  output$history_table <- renderTable({
    req(counter() > 0)
    history_data()
  }, striped = TRUE, hover = TRUE, spacing = "m", align = "c", width = "100%")
  
  # Generating PDF Reports
  output$download_report <- downloadHandler(
    filename = function() {
      paste0("Stroke_Risk_Report_", Sys.Date(), ".pdf")
    },
    content = function(file) {
      req(counter() > 0)
      tempReport <- file.path(tempdir(), "report.Rmd")
      
      report_text <- paste0(
        "---\n",
        "title: \"Stroke Assessment Clinical Report\"\n",
        "date: \"", Sys.Date(), "\"\n",
        "output: pdf_document\n",
        "---\n\n",
        "### 1. Patient Demographics & Background Clinical Metrics\n",
        "***\n",
        "- **Age:** ", input$age, " years old\n",
        "- **Gender:** ", input$gender, "\n",
        "- **Hypertension History:** ", ifelse(input$hypertension == "1", "Yes", "No"), "\n",
        "- **Heart Disease History:** ", ifelse(input$heart_disease == "1", "Yes", "No"), "\n",
        "- **Average Glucose Level:** ", input$glucose, " mg/dL\n",
        "- **Body Mass Index (BMI):** ", input$bmi, " kg/m²\n\n",
        "### 2. Predictive Diagnostics Summary Output\n",
        "***\n",
        "- **Probability of having a Stroke:** **", results$stroke_p, "%**\n",
        "- **Probability of NOT having a Stroke:** **", results$no_stroke_p, "%**\n\n",
        "### 3. Clinical Disclaimer\n",
        "***\n",
        "*This analytical summary report is generated purely for screening support, automated risk calculation, and scientific evaluations.*"
      )
      
      writeLines(report_text, con = tempReport)
      rmarkdown::render(tempReport, output_file = file, envir = new.env(parent = globalenv()))
    }
  )
}

# =====================
# RUN APP
# =====================
shinyApp(ui, server)

saveRDS(final_model, "model.rds")
getwd()


# install.packages("rsconnect")
# library(rsconnect)
# 
# 
# 
# rsconnect::setAccountInfo(name='udara',
# 			  token='5DA3E49B0485AC0295058589D3D83D6F',
# 			  secret='')
```
``




# Task Five: Findings and Conclusions
































